---
output: github_document
---

```{r intro, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
library(readr)
library(dplyr)
library(tidyr)
library(maps)
library(magrittr)
library(Fars)

```

# Fars package overview

The Fars package contains a series of functions that enable the user to quickly load and summarize motor vehicle crash data. 

The functions contained within this package are as follows:

* fars_read
* make_filename
* fars_read_years
* fars_summarize_years
* fars_map_state

## fars_read

The fars_read function enables users to read motor vehicle crash data from a csv file.
The input to the function is the path, name, and extension of the csv file containing the data.
The output is a data frame. 
The examples below use the 2013 and 2014 motor vehicle crash data that is included with this package and can be accessed using the following codes:

* system.file("extdata", "accident_2013.csv.bz2", package = "Fars")
* system.file("extdata", "accident_2014.csv.bz2", package = "Fars").
 
Examples:
```{r read_ex}
x1 <- system.file("extdata", "accident_2013.csv.bz2", package = "Fars")
y1 <- fars_read(x1)
y1

x2 <- system.file("extdata", "accident_2014.csv.bz2", package = "Fars")
y2 <- fars_read(x2)
y2
```

## make_filename

The make_filename function enables users to generate the csv filename for a particular year of motor vehicle crash data.
The input to the function is a year.
The output is a file name and extension. 
 
Examples:
```{r make_ex}
x1 <- make_filename(2013)
x1

x2 <- make_filename(2014)
x2
```

## fars_read_years

The fars_read_years function enables users to read motor vehicle crash data from a csv file for a particular year (or list/vector of years).
The data must be stored in the current directory.
The output is a data frame with only the month and year columns. 
 
Examples:
```{r read_year_ex}
x1 <- fars_read_years(2013)
x1

x2 <- fars_read_years(c(2013, 2014))
x2
```

## fars_summarize_years

The fars_summarize_years function enables users to read motor vehicle crash data from a csv file for a particular year (or list/vector of years).
The data must be stored in the current directory.
The output is a data frame with only the month and year columns.
The number of motor vehicle crashes are represented with months displayed along the rows and years along the columns. 
 
Examples:
```{r summarize_year_ex}
x1 <- fars_summarize_years(2013)
x1

x2 <- fars_summarize_years(c(2013, 2014))
x2
```

## fars_map_state

The fars_map_state function enables users to read motor vehicle crash data from a csv file for a particular year and state number.
The data must be stored in the current directory.
The output is a graph showing accident locations in a particular year on a map of the user-defined state.
 
Examples:
```{r map_state_ex}
fars_map_state(1, 2013)
fars_map_state(48, 2014)
```
